Models

### 2021-06-30 analysis/05_model_metrics.Rmd
Models of study-level slope estimates. Minor changes to figures.

metric_model_table <- bind_rows(SES_rand_slope, s_mod, jacc_base_mod)
```

@@

Taxa:
```{r}
taxa_slope_model <- model_data %>%
  filter(metric %in%   c("SES_FRic", "SES_FEve", "SES_FDiv")) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(value ~ year_scaled + taxa + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
                                 data = .x, REML = TRUE)),
           coef = map(model, tidy)) %>%
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  #filter(term == "year_scaled") %>%
  filter(term %in% c("year_scaled", "taxaBirds", "taxaMammals")) %>%
  select(-c(effect, group, statistic))

taxa_slope_model_S <- model_data %>% 
  mutate(year_scaled = scale(year)) %>% 
  filter(metric == "S") %>%
  lmer(logvalue ~ year_scaled + taxa + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
                                 data = ., REML = TRUE)
```
Realm:
```{r}
realm_slope_model <- model_data %>%
  filter(metric %in%   c("SES_FRic", "SES_FEve", "SES_FDiv")) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(value ~ year_scaled + realm + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
                                 data = .x, REML = TRUE)),
           coef = map(model, tidy)) %>%
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  #filter(term == "year_scaled") %>%
  filter(effect == "fixed") %>%
  select(-c(effect, group, statistic))

@@

#### Analizing slopes
Pairwise correlation analysis
```{r}
library(psych)
#slope_corr <- corr.test(slopes_wide %>% ungroup() %>% select(starts_with("SES"), "S", "Jaccard_base"))

@@

  trend_plot <- data %>%
    left_join(pred, by = "year_scaled") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = !!sym(yval))) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.50) +
    geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme_classic() +
    theme(axis.title = element_text(size=7), legend.position = "none") +
    ylab(ylabel) +
    xlab(xlabel) +
    labs(color = "Climate") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
    scale_colour_manual(values = color_list, drop = FALSE)

@@

######
group.colors <- c(A = "#333BFF", B = "#CC6600", C ="#9633FF", D = "#E2FF33", E = "#E3DB71")

simplePlot <- function(DT, tit) 
  ggplot(DT ,aes(x=Name, y=Value, fill=Group)) +
  geom_bar(stat="identity") + xlab("") + ggtitle(tit) +
  #Specify colours
  scale_fill_manual(values=group.colors)
#####

@@

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
plot_join + draw_grob(legend,  2/3.3, -0.3, .3/3.3, 1, scale = 0.9)

>>>>

metric_model_table <- bind_rows(SES_rand_slope, s_mod, jacc_base_mod)

#save coefficients and models
usethis::use_data(metric_model_table)
pins::pin(metric_model_table, board = "github")
```

@@

Taxa:
```{r}
# taxa_slope_model <- model_data %>%
#   filter(metric %in%   c("SES_FRic", "SES_FEve", "SES_FDiv")) %>%
#   group_nest(metric) %>%
#   mutate(model = map(data, ~lmer(value ~ year_scaled + taxa + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
#                                  data = .x, REML = TRUE)),
#            coef = map(model, tidy)) %>%
#   select(metric, coef) %>%
#   unnest(cols = c(coef)) %>%
#   #filter(term == "year_scaled") %>%
#   filter(term %in% c("year_scaled", "taxaBirds", "taxaMammals")) %>%
#   select(-c(effect, group, statistic))
# 
# taxa_slope_model_S <- model_data %>% 
#   mutate(year_scaled = scale(year)) %>% 
#   filter(metric == "S") %>%
#   lmer(logvalue ~ year_scaled + taxa + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
#                                  data = ., REML = TRUE)

SES_rand_slope <- model_data %>%
  filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         model = map(data, ~lmer(value ~ taxa + (1|study_id) + (|study_id:rarefyID), 
                                 data = .x)),
           # null_model = map(data, ~lmer(value ~ (year_scaled|study_id/rarefyID), 
           #                       data = .x)),
           # test = map2(model, null_model, anova),
           # test_res = map(test, tidy),
           coef = map(model, tidy),
           ) %>% 
  select(metric, model, coef) %>%
  unnest(cols = c(coef)) %>%
  filter(term == "year_scaled") %>%
  select(-c(effect, group, term, statistic))
```
Realm:
```{r}
realm_slope_model <- model_data %>%
  filter(metric %in%   c("SES_FRic", "SES_FEve", "SES_FDiv")) %>%
  group_nest(metric) %>%
  mutate(model = map(data, ~lmer(value ~ year_scaled + realm + (year_scaled|study_id/rarefyID), # + (year_scaled|taxa), 
                                 data = .x, REML = TRUE)),
           coef = map(model, tidy)) %>%
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  #filter(term == "year_scaled") %>%
  filter(effect == "fixed") %>%
  select(-c(effect, group, statistic))

@@

#### Analizing slopes
Is slope a function of timeseries characteristics?
```{r}
study_slope_data <- study_slopes %>%
  select(study_id, metric, estimate) %>%
  left_join(meta %>% 
              select(-rarefyid) %>% 
              distinct() %>%
              mutate(study_id = as.character(study_id))) %>%
  left_join(model_data %>% select(study_id, startyear, duration) %>% distinct())

#taxa model
taxa_slope <- study_slope_data %>%
  #filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         #model = map(data, ~lmer(estimate ~ taxa + (1|taxa), 
         model = map(data, ~glm(estimate ~ taxa, 
                                 data = .x)),
           coef = map(model, tidy),
           ) %>% 
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  mutate(covariate = "taxa")

climate_slope <- study_slope_data %>%
  #filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         #model = map(data, ~lmer(estimate ~ taxa + (1|taxa), 
         model = map(data, ~glm(estimate ~ climate, 
                                 data = .x)),
           coef = map(model, tidy),
           ) %>% 
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  mutate(covariate = "climate")

realm_slope <- study_slope_data %>%
  #filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         #model = map(data, ~lmer(estimate ~ taxa + (1|taxa), 
         model = map(data, ~glm(estimate ~ realm, 
                                 data = .x)),
           coef = map(model, tidy),
           ) %>% 
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  mutate(covariate = "realm")

duration_slope <- study_slope_data %>%
  #filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         #model = map(data, ~lmer(estimate ~ taxa + (1|taxa), 
         model = map(data, ~glm(estimate ~ duration, 
                                 data = .x)),
           coef = map(model, tidy),
           ) %>% 
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  mutate(covariate = "duration")

startyear_slope <- study_slope_data %>%
  #filter(metric %in%  c("SES_FRic", "SES_FEve", "SES_FDiv", "Jaccard_next")) %>%
  group_nest(metric) %>%
  mutate(#year_center = map(data, ~scale(.x$year)),
         #model = map(data, ~lmer(estimate ~ taxa + (1|taxa), 
         model = map(data, ~glm(estimate ~ startyear, 
                                 data = .x)),
           coef = map(model, tidy),
           ) %>% 
  select(metric, coef) %>%
  unnest(cols = c(coef)) %>%
  mutate(covariate = "startyear")

#All model estimates in one dataframe
slope_models <- bind_rows(taxa_slope, realm_slope, climate_slope, duration_slope, startyear_slope)

# Which predictor variables were significant?
slope_models %>% filter(term != "(Intercept)", p.value < 0.05) %>% View()
```

Pairwise correlation analysis
```{r}
library(psych)
#slope_corr <- corr.test(slopes_wide %>% ungroup() %>% select(starts_with("SES"), "S", "Jaccard_base"))

@@

  trend_plot <- data %>%
    left_join(pred, by = "year_scaled") %>%
    #left_join(metadata %>% select(study_id, climate)) %>%
    ggplot(aes(x = year, y = !!sym(yval))) +
    # geom_smooth(aes(color = climate, group = rarefyID), method = "glm", se = FALSE) +
    # geom_point(aes(color = climate, group = rarefyID)) +
    geom_point(aes(group = rarefyID) ,color = "grey", size = 0.25) +
    geom_smooth(aes(color = climate, group = study_id), method = "glm", se = FALSE, size = 0.7) +
    #geom_smooth(color = "black", method = "glm", se = FALSE) +
    geom_line(mapping = aes(y = pred)) +
    theme_classic() +
    theme(axis.title = element_text(size=7), legend.position = "none") +
    ylab(ylabel) +
    xlab(xlabel) +
    labs(color = "Climate") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6)) +
    scale_colour_manual(values = color_list, drop = FALSE)

@@

@@

# add the legend to the row we made earlier. Give it one-third of 
# the width of one plot (via rel_widths).
#plot_join + draw_grob(legend,  2/3.3, -0.3, .3/3.3, 1, scale = 0.6)
plot_join + draw_grob(legend,  2/3.3, -0.3,.3/3.3, 1, scale = 0.01)
